### 2024-12-19 manuscript/S1_equations.tex
Create S1_equations.tex

Let $m^*$ be mean crowding and $m$ be the mean. By definition, we have $m^* = m +(\sigma^2/m -1)$. 
Therefore, we have that $m^*- m = \sigma^2/m -1$.

We also have that $\sigma^2 = m + m^2/\theta$, by definition of negative binomial variance, where $\theta$ is the dispersion parameter. This is equivalent to $\sigma^2/m = 1 + m/\theta$.
Therefore, we have that $m/\theta = \sigma^2/m - 1$.

By equating these, we have $m^*- m = m/\theta$.

Dividing both sides by $m$, we have $m^*/m - 1 = 1/\theta$

Finally, we have that $m^*= m/\theta + m$