### 2025-02-15 ms-3/S1_text.tex
Update S1_text.tex

\documentclass{article}
\usepackage{amsmath}
\begin{document}

{\noindent \bf S1 Text. Derivation of the relationship between the disperion parameter and the mean crowding parameter.}\\

\noindent 1. Let $m^*$ be mean crowding and $m$ be the mean.\\
$m^* = m +(\frac{\sigma^2}{m} -1)$ (by definition of mean crowding)\\
$ \implies m^*- m = \frac{\sigma^2}{m} -1$\\

\noindent 2. Let $\theta$ be the dispersion parameter and $\sigma^2$ be the variance of the negative binomial distribution.\\ 
$\sigma^2 = m + \frac{m^2}{\theta}$ (by definition of negative binomial variance)\\ $\implies \frac{\sigma^2}{m} = 1 + \frac{m}{\theta}$\\
$\implies \frac{m}{\theta} = \frac{\sigma^2}{m} - 1$\\

\noindent 3. $\frac{m^*}{m} - 1 = \frac{1}{\theta}$ (by equating the left-hand sides of the results in 1 and 2 and dividing by $m$) \\
$\implies m^*= m + \frac{m}{\theta}$

\end{document}